---
title: "API to Oracle Example"
author: "Dykema, Bill"
date: "2021/10/01"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### install.packages("rmarkdown")
#### Knit to HTML for easy viewing: rmarkdown::render("api-oracle.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# API to Oracle example

## Load R libraries
```{r libraries}

# GET request packages for reading from the API
#install.packages("httr")
library(httr)

# JSON converter
#install.packages("jsonlite")
library(jsonlite)

# Data analysis packages
#install.packages("tidyverse")
library(tidyverse)

# ODBC package used for Oracle
# install.packages("RODBC")
library(RODBC)

```

## API Connect and Pull Data
```{r read_api}

# API URL build
drought_area <- "USStatistics"
drought_type <- "GetDroughtSeverityStatisticsByArea"
drought_aoi <- "us"
drought_start <- "8/10/2021"
drought_end <- "8/20/2021"
drought_stat <- "1"
drought_url <- sprintf("https://usdmdataservices.unl.edu/api/%s/%s?aoi=%s&startdate=%s&enddate=%s&statisticsType=%s", drought_area, drought_type, drought_aoi, drought_start, drought_end, drought_stat)
# Send API request
drought_data <- GET(drought_url)
# Check URL status (200 is good)
drought_data
# Convert to DataFrame
df_drought <- fromJSON(content(drought_data, "text"), flatten=TRUE)
# View new DataFrame
df_drought

```

## API Columns
```{r api_columns}

# Display DataFrame columns
summary(df_drought)
str(df_drought)

```

# Format Fields
```{r api_columns_setup}

# Copy DataFrame
df_drought_transformed <- df_drought

# Character to Date/Time
df_drought_transformed[['MapDate']] <- parse_date(df_drought[['MapDate']], '%Y%m%d')
df_drought_transformed[['ValidStart']] <- parse_date(df_drought[['ValidStart']], '%Y-%m-%d')
df_drought_transformed[['ValidEnd']] <- parse_date(df_drought[['ValidEnd']], '%Y-%m-%d')
# Remove commas in field values
df_drought_transformed[['None']] <- as.numeric(gsub(",","",df_drought[['None']]))
df_drought_transformed[['D0']] <- as.numeric(gsub(",","",df_drought[['D0']]))
df_drought_transformed[['D1']] <- as.numeric(gsub(",","",df_drought[['D1']]))
df_drought_transformed[['D2']] <- as.numeric(gsub(",","",df_drought[['D2']]))
df_drought_transformed[['D3']] <- as.numeric(gsub(",","",df_drought[['D3']]))
df_drought_transformed[['D4']] <- as.numeric(gsub(",","",df_drought[['D4']]))
# View DataFrame
df_drought_transformed

# Display default may appear to cut-off decimals, you can override this
#options(digits=20)
#df_drought_transformed

```

## Transformed Columns
```{r transformed_columns}

# Display DataFrame columns
summary(df_drought_transformed)
str(df_drought_transformed)

```

## Add Record Added Field
```{r transformed_record_added}

## Get the current date/time
now_is <- Sys.time()

## Populate the updated date/time
df_drought_transformed[['record_added']] <- now_is

```

## Oracle Setup and Test
```{r oracle_setup_test}

# Connection string (assumes Oracle ODBC DSN is already configured)
odbc_dsn <- "DSNNameHere"

# Connect to DSN string
oracle_odbc <- odbcConnect(odbc_dsn)

# Table to use for test
test_table <- "apitable"

# Query to use for test
test_query <- sprintf("SELECT * FROM %s", test_table)

# Run test query
df_oracle <- sqlQuery(oracle_odbc, test_query)

# Display results of test query
df_oracle

# Display table column types (only needed for sqlSave)
oracle_columns <- sqlColumns(oracle_odbc, test_table)

```

## Optional Drop Table before Loading Data
```{r oracle_drop}

# Build statement
oracle_drop <- sprintf("DROP TABLE %s", test_table)

# Drop table
sqlQuery(oracle_odbc, oracle_drop)

# Build statement to recreate empty table
oracle_create <- sprintf("CREATE TABLE %s (MapDate DATE NOT NULL, AreaOfInterest VARCHAR2(50), None NUMBER(10,2), D0 NUMBER(10,2), D1 NUMBER(10,2), D2 NUMBER(10,2), D3 NUMBER(10,2), D4 NUMBER(10,2), ValidStart DATE, ValidEnd DATE, StatisticFormatID VARCHAR2(50), record_added TIMESTAMP)", test_table)
# Create table
sqlQuery(oracle_odbc, oracle_create)

```

## Append API data to Oracle
```{r api_oracle_append}

# Append API DataFrame to Oracle table
for (dataframe_row in rownames(df_drought_transformed)) {
  # Build SQL insert
  oracle_record_add <- sprintf("INSERT INTO %s (MapDate, AreaOfInterest, None, D0, D1, D2, D3, D4, ValidStart, ValidEnd, StatisticFormatID, record_added) VALUES (TO_DATE('%s', 'YYYY-MM-DD'), '%s', '%d', '%d', '%d', '%d', '%d', '%d', TO_DATE('%s', 'YYYY-MM-DD'), TO_DATE('%s', 'YYYY-MM-DD'), '%s')", test_table, df_drought_transformed[dataframe_row, "MapDate"], df_drought_transformed[dataframe_row, "AreaOfInterest"], df_drought_transformed[dataframe_row, "None"], df_drought_transformed[dataframe_row, "D0"], df_drought_transformed[dataframe_row, "D1"], df_drought_transformed[dataframe_row, "D2"], df_drought_transformed[dataframe_row, "D3"], df_drought_transformed[dataframe_row, "D4"], df_drought_transformed[dataframe_row, "ValidStart"], df_drought_transformed[dataframe_row, "ValidEnd"], df_drought_transformed[dataframe_row, "StatisticFormatID"], df_drought_transformed[dataframe_row, "record_added"])
  # Run SQL insert
  sqlQuery(oracle_odbc, oracle_record_add)
}

```

## Oracle Test for New Data
```{r oracle_data_test}

# Run test query
df_oracle_new <- sqlQuery(oracle_odbc, test_query)

# Display results of test query
df_oracle_new

```

## Oracle Close Connection
```{r oracle_close}

# Close the Oracle connection
close(oracle_odbc)

```