---
title: "Excel to/from CSV Example"
author: "Dykema, Bill"
date: "2021/09/09"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### install.packages("rmarkdown")
#### Knit to HTML for easy viewing: rmarkdown::render("excel-csv.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Excel and CSV read and write example

## Load R libraries

```{r libraries}

# Data analysis packages
#install.packages("tidyverse")
library(tidyverse)

# Excel packages
#install.packages("xlsx")
library(xlsx)

```

## Read CSV file

```{r read_csv_file}

# Create DataFrame from CSV file
df_csv <- read.csv("testtable_source.csv", stringsAsFactors=FALSE)

# View your new DataFrame
df_csv

```

## Read Excel file

```{r read_excel_file}

# Create DataFrame from Excel file
df_excel <- read.xlsx2("testtable_source.xlsx", sheetIndex=1)

# View your new DataFrame
df_excel

```

## CSV Columns

```{r csv_columns}

# Display DataFrame columns
summary(df_csv)
str(df_csv)

```

## Excel Columns

```{r excel_columns}

# Display DataFrame columns
summary(df_excel)
str(df_excel)

```

## CSV Setup Columns

```{r csv_columns_setup}

# Rename columns
df_csv %>%
  rename(test_datetime = TEST_DATETIME) %>%
  rename(test_number = TEST_NUMBER) %>%
  rename(test_text = TEST_TEXT) %>%
  rename(record_added = RECORD_ADDED) ->
df_csv_transformed

```

## Excel Setup Columns

```{r excel_columns_setup}

# Rename columns
df_excel %>%
  rename(test_datetime = TEST_DATETIME) %>%
  rename(test_number = TEST_NUMBER) %>%
  rename(test_text = TEST_TEXT) %>%
  rename(record_added = RECORD_ADDED) ->
df_excel_transformed

# Format date fields in DataFrame
df_excel_transformed[['test_datetime']] <- openxlsx::convertToDateTime(df_excel[['TEST_DATETIME']])
df_excel_transformed[['record_added']] <- openxlsx::convertToDateTime(df_excel[['RECORD_ADDED']])

```

## Write CSV to Excel
```{r csv_to_excel}

# Send CSV DataFrame to Excel file
write.xlsx2(df_csv_transformed, file='testtable_csv_export.xlsx', row.names = FALSE)

```

## Write Excel to CSV
```{r excel_to_csv}

# Send Excel DataFrame to CSV file
write.csv(df_excel_transformed, file='testtable_csv_export.csv', row.names = FALSE)

```

## Excel/CSV Oldest and Latest Dates
```{r excel_csv_oldlate}

# Get oldest CSV date
date_oldest_csv <- min(df_csv_transformed[['test_datetime']])
cat (c("Oldest CSV date is...\n"))
date_oldest_csv
# Get newest Excel date
date_newest_excel <- max(df_excel_transformed[['test_datetime']])
cat (c("Newest Excel date is...\n"))
date_newest_excel

```

## Write CSV transform to Excel
```{r csv_transform}

# Filter the CSV data
df_csv_filtered <- df_csv_transformed[df_csv_transformed$test_datetime==date_oldest_csv,]
# Select specific columns
df_csv_filtered %>%
  select(test_datetime, test_number) ->
df_csv_filtered
# Send part part of the CSV filtered DataFrame to Excel file
write.xlsx2(df_csv_filtered, file='testtable_transformed_csv_export.xlsx', row.names = FALSE)

```

## Write Excel transform to CSV
```{r excel_transform}

# Filter the Excel data
df_excel_filtered <- df_excel_transformed[df_excel_transformed$test_datetime==date_newest_excel,]
# Select specific columns
df_excel_filtered %>%
  select(test_datetime, test_number) ->
df_excel_filtered
# Send part part of the Excel filtered DataFrame to CSV file
write.csv(df_excel_filtered, file='testtable_transformed_excel_export.csv', row.names = FALSE)

```

## Excel/CSV Combine into Single Excel/CSV
```{r excel_csv_combine}

# Append two DataFrames
df_excel_combined <- rbind(df_excel_transformed, df_csv_transformed)

# View combined DataFrame
df_excel_combined

# Send combined DataFrame to Excel and CSV file
write.xlsx2(df_excel_combined, file='testtable_combined_export.xlsx', row.names = FALSE)
write.csv(df_excel_combined, file='testtable_combined_export.csv', row.names = FALSE)

```

## Write Two Seperate Data Sets to Excel in Seperate Sheets
```{r excel_sheets}

# Send date to Excel sheets
excel_file <- "testtable_sheets_export.xlsx"
write.xlsx2(df_csv_filtered, file=excel_file, sheetName="CSVData", row.names = FALSE)
write.xlsx2(df_excel_filtered, file=excel_file, sheetName="ExcelData", row.names = FALSE, append=TRUE)

```