---
title: "Oracle to Oracle Example"
author: "Dykema, Bill"
date: "2021/09/23"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### install.packages("rmarkdown")
#### Knit to HTML for easy viewing: rmarkdown::render("oracle-oracle.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Oracle to Oracle example

## Load R libraries
```{r libraries}

# ODBC package used for Oracle
# install.packages("RODBC")
library(RODBC)

# Data analysis packages
#install.packages("tidyverse")
library(tidyverse)

```

## Oracle Setup
```{r oracle_setup}

# Connection string (assumes Oracle ODBC DSN is already configured)
odbc_dsn <- "DSNNameHere"

# Connect to DSN string
oracle_odbc <- odbcConnect(odbc_dsn)

```

## Read Oracle Table
```{r oracle_read}

# Table to use
oracle_table <- "testtable"

# Query to use
oracle_query <- sprintf("SELECT * FROM %s", oracle_table)

# Run query
df_oracle <- sqlQuery(oracle_odbc, oracle_query)
# Alternatively you can use this, but it might not be compatible with all databases
#df_oracle <- sqlFetch(oracle_odbc, oracle_table)

# Display results of test query
df_oracle

# Display table column types (only needed for sqlSave)
oracle_columns <- sqlColumns(oracle_odbc, oracle_table)

```

## Oracle Setup Columns
```{r excel_columns_setup}

# Rename columns
df_oracle %>%
  rename(test_datetime = TEST_DATETIME) %>%
  rename(test_number = TEST_NUMBER) %>%
  rename(test_text = TEST_TEXT) %>%
  rename(record_added = RECORD_ADDED) ->
df_oracle_transformed

```

## Oracle Columns
```{r oracle_columns}

# Display DataFrame columns
summary(df_oracle_transformed)
str(df_oracle_transformed)

```

## Update Record Added Fields
```{r oracle_updated}

## Get the current date/time
now_is <- Sys.time()

## Populate the updated date/time
df_oracle_transformed[['record_added']] <- now_is

```

## Optional Drop Table before Loading Data
```{r oracle_drop}

# Build statement
oracle_drop <- sprintf("DROP TABLE %s", oracle_table)

# Drop table
sqlQuery(oracle_odbc, oracle_drop)
# Alternatively you can use this
#sqlDrop(oracle_odbc, oracle_table, errors=TRUE)

# Build statement to recreate empty table
oracle_create <- sprintf("CREATE TABLE %s (test_datetime TIMESTAMP NOT NULL, test_number INT, test_text VARCHAR2(50), record_added TIMESTAMP)", oracle_table)
# Create table
sqlQuery(oracle_odbc, oracle_create)

```

## Set Field types
```{r oracle_fields}

# Set field types used in the Oracle table, only needed if using sqlSave
oracle_field_types <- as.character(as.character(c("varchar2(255)", "varchar2(255)","int(38)","varchar2(255)","int(38)")))
names(oracle_field_types) <- as.character(oracle_columns$COLUMN_NAME)

# Display Oracle field types to use if using sqlSave
oracle_field_types

```

## Append Oracle to Oracle
```{r oracle_append}

# Append Oracle DataFrame to Oracle table
for (dataframe_row in rownames(df_oracle_transformed)) {
  # Build SQL insert
  oracle_record_add <- sprintf("INSERT INTO %s (test_datetime, test_number, test_text, record_added) VALUES (TO_DATE('%s', 'YYYY/MM/DD HH24:MI:SS'), '%s', '%s', TO_DATE('%s', 'YYYY/MM/DD HH24:MI:SS'))", oracle_table, df_oracle_transformed[dataframe_row, "test_datetime"], df_oracle_transformed[dataframe_row, "test_number"], df_oracle_transformed[dataframe_row, "test_text"], df_oracle_transformed[dataframe_row, "record_added"])
  # Run SQL insert
  sqlQuery(oracle_odbc, oracle_record_add)
}

# Alternatively you can use this rather than a loop, but it can be slower and not compatible with all databases
#sqlSave(oracle_odbc, df_oracle_transformed, oracle_table, rownames=TRUE, append=TRUE, varTypes=oracle_field_types, verbose=TRUE)

```

## Oracle Test for New Data
```{r oracle_data_test}

# Run test query
df_oracle_new <- sqlQuery(oracle_odbc, oracle_query)

# Display results of test query
df_oracle_new

```

## Oracle Oldest and Latest Dates
```{r oracle_oldlate}

# Get oldest Oracle date
date_oldest <- min(df_oracle_transformed[['test_datetime']])
print("Oldest Oracle date is...")
date_oldest

```

## Write Oracle transform to Oracle
```{r oracle_transform_oracle}

# Filter the Oracle data
df_oracle_filtered <- df_oracle_transformed[df_oracle_transformed$test_datetime==date_oldest,]

# Append filtered Oracle DataFrame to Oracle table
for (dataframe_row in rownames(df_oracle_filtered)) {
  # Build SQL insert
  oracle_record_add <- sprintf("INSERT INTO %s (test_datetime, test_number, test_text, record_added) VALUES (TO_DATE('%s', 'YYYY/MM/DD HH24:MI:SS'), '%s', '%s', TO_DATE('%s', 'YYYY/MM/DD HH24:MI:SS'))", oracle_table, df_oracle_filtered[dataframe_row, "test_datetime"], df_oracle_filtered[dataframe_row, "test_number"], df_oracle_filtered[dataframe_row, "test_text"], df_oracle_filtered[dataframe_row, "record_added"])
  # Run SQL insert
  sqlQuery(oracle_odbc, oracle_record_add)
}

# Alternatively you can use this rather than a loop, but it can be slower and not compatible with all databases
#sqlSave(oracle_odbc, df_oracle_filtered, oracle_table, rownames=TRUE, append=TRUE, varTypes=oracle_field_types, verbose=TRUE)

```

## Oracle Test for Appended Data
```{r oracle_transform_appended}

# Run test query
df_oracle_appended <- sqlQuery(oracle_odbc, oracle_query)

# Display results of test query
df_oracle_appended

```

## Oracle Close Connection
```{r oracle_close}

# Close the Oracle connection
close(oracle_odbc)

```