---
title: "Oracle to Excel/CSV Example"
author: "Dykema, Bill"
date: "2021/09/16"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### install.packages("rmarkdown")
#### Knit to HTML for easy viewing: rmarkdown::render("oracle-excel.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Oracle to Excel/CSV example

## Load R libraries
```{r libraries}

# ODBC package used for Oracle
# install.packages("RODBC")
library(RODBC)

# Data analysis packages
#install.packages("tidyverse")
library(tidyverse)

# Excel packages
#install.packages("xlsx")
library(xlsx)
#install.packages("openxlsx")
library(openxlsx)

```

## Oracle Setup
```{r oracle_setup}

# Connection string (assumes Oracle ODBC DSN is already configured)
odbc_dsn <- "DSNNameHere"

# Connect to DSN string
oracle_odbc <- odbcConnect(odbc_dsn)

```

## Read Oracle Table
```{r oracle_read}

# Table to use
oracle_table <- "testtable"

# Query to use
oracle_query <- sprintf("SELECT * FROM %s", oracle_table)

# Run query
df_oracle <- sqlQuery(oracle_odbc, oracle_query)
# Alternatively you can use this, but it might not be compatible with all databases
#df_oracle <- sqlFetch(oracle_odbc, oracle_table)

# Display results of test query
df_oracle

# Display table column types (only needed for sqlSave)
oracle_columns <- sqlColumns(oracle_odbc, oracle_table)

```

## Oracle Close Connection
```{r oracle_close}

# Close the Oracle connection
close(oracle_odbc)

```

## Oracle Setup Columns
```{r excel_columns_setup}

# Rename columns
df_oracle %>%
  rename(test_datetime = TEST_DATETIME) %>%
  rename(test_number = TEST_NUMBER) %>%
  rename(test_text = TEST_TEXT) %>%
  rename(record_added = RECORD_ADDED) ->
df_oracle_transformed

```

## Oracle Columns
```{r oracle_columns}

# Display DataFrame columns
summary(df_oracle_transformed)
str(df_oracle_transformed)

```

## Write Oracle to Excel
```{r oracle_to_excel}

# Send Oracle DataFrame to Excel file
write.xlsx2(df_oracle_transformed, file='testtable_oracle_export.xlsx', row.names=FALSE)

```

## Write Oracle to CSV
```{r oracle_to_csv}

# Send Oracle DataFrame to CSV file
write.csv(df_oracle_transformed, file='testtable_oracle_export.csv', row.names=FALSE)

```

## Write Oracle to Text File with some Text Manipulation
```{r oracle_to_file}

# File name
file_out <- "testtable_oracle_export.txt"

# Filter header
write("This is a test file!", file_out)

# DataFrame to Text Formatted
for (dataframe_row in rownames(df_oracle_transformed)) {
  # Build formatted text line
  text_record_add <- sprintf("On %s we inserted record %d, details are '%s', original date was %s", df_oracle_transformed[dataframe_row, "record_added"], df_oracle_transformed[dataframe_row, "test_number"], df_oracle_transformed[dataframe_row, "test_text"], df_oracle_transformed[dataframe_row, "test_datetime"])
  # Send Line of Text to file
  write(text_record_add, file_out, append=TRUE)
}

```

## Oracle Oldest and Latest Dates
```{r oracle_oldlate}

# Get oldest CSV date
date_oldest <- min(df_oracle_transformed[['test_datetime']])
print("Oldest Oracle date is...")
date_oldest
# Get newest Excel date
date_newest <- max(df_oracle_transformed[['test_datetime']])
print("Newest Oracle date is...")
date_newest

```

## Write Oracle transform to Excel
```{r oracle_transform_excel}

# Filter the Oracle data
df_oracle_filtered_excel <- df_oracle_transformed[df_oracle_transformed$test_datetime==date_oldest,]
# Select specific columns
df_oracle_filtered_excel %>%
  select(test_datetime, test_number) ->
df_oracle_filtered_excel
# Send part part of the filtered DataFrame to Excel file
write.xlsx2(df_oracle_filtered_excel, file='testtable_oracle_transformed_export.xlsx', row.names=FALSE)

```

## Write Oracle transform to CSV
```{r oracle_transform_csv}

# Filter the Excel data
df_oracle_filtered_csv <- df_oracle_transformed[df_oracle_transformed$test_datetime==date_newest,]
# Select specific columns
df_oracle_filtered_csv %>%
  select(test_datetime, test_number) ->
df_oracle_filtered_csv
# Send part part of the filtered DataFrame to CSV file
write.csv(df_oracle_filtered_csv, file='testtable_oracle_transformed_export.csv', row.names=FALSE)

```

## Write Two Seperate Data Sets to Excel in Seperate Sheets
```{r oracle_excel_sheets}

# Send date to Excel sheets
excel_sheets_file <- "testtable_oracle_sheets_export.xlsx"
write.xlsx2(df_oracle_filtered_excel, file=excel_sheets_file, sheetName="ExcelData", row.names=FALSE)
write.xlsx2(df_oracle_filtered_csv, file=excel_sheets_file, sheetName="CSVData", row.names=FALSE, append=TRUE)

```