---
title: "ODBC DSN Set Using External R Script"
author: "Dykema, Bill"
date: "2021/09/08"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### install.packages("rmarkdown")
#### Knit to HTML for easy viewing: rmarkdown::render("roracle_odbc-select.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ODBC DSN by Folder Location Example Using External R Script
#### This script shows how to set the Oracle ODBC DSN by running an external script

# Get Oracle ODBC DSN using external R script

```{r external_script}

source("roracle_odbc-select.R", local = knitr::knit_global())

```

# Load R libraries

```{r libraries}

## Load libraries

# install.packages("RODBC")
library(RODBC)

```

## DSN test

```{r dsn_test}

# Table to use for test
test_table <- "testtable"

# Query to use for test
test_query <- sprintf("SELECT * FROM %s", test_table)

# Connect to DSN string we were just given from environmental variable based on folder
oracle_odbc <- odbcConnect(odbc_dsn)

## Run test query
records_test <- sqlQuery(oracle_odbc, test_query)

## Display results of test query
records_test

```
