---
title: "ODBC DSN by Folder Location"
author: "Dykema, Bill"
date: "2021/09/08"
output:
  html_document: default
---

## Notes for exporting to HTML file:
#### Set working directory: setwd("[Working Folder]")
#### Knit to HTML for easy viewing: rmarkdown::render("roracle_odbc-select.Rmd", "all")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ODBC DSN by Folder Location Example
#### This script shows how to set the Oracle ODBC DSN based on its current folder location

# Load R libraries

```{r libraries}

## Load libraries

#install.packages("RODBC")
library(RODBC)

```

# Read server environmental variables for the Oracle ODBC DSN in each environment

```{r variables}

# Environments
env_dsn_dev <- Sys.getenv("R_DSN_DEV")    ## Server environmental variable name for development Oracle ODBC DSN name
env_dsn_crt <- Sys.getenv("R_DSN_CRT")    ## Server environmental variable name for test Oracle ODBC DSN name
env_dsn_prod <- Sys.getenv("R_DSN_PROD")  ## Server environmental variable name for production Oracle ODBC DSN name

```

## Determine current folder location

```{r folder_location}

# Current folder
env_folder <- getwd()

```

## DSN set based on folder

```{r dsn_from_folder}

# Set Oracle ODBC DSN based on folder
if (grepl("/PROD", env_folder, fixed = TRUE)) {       ## Contains "/PROD" in the folder path
  odbc_dsn <- env_dsn_prod
} else if (grepl("/CRT", env_folder, fixed = TRUE)) { ## Contains "/CRT" in the folder path
  odbc_dsn <- env_dsn_crt
} else {                                              ## Does not contain "/PROD" or "/CRT" in the folder path, must be Dev
  odbc_dsn <- env_dsn_dev
}

```

## DSN test

```{r dsn_test}

# Table to use for test
test_table = "testtable"

# Query to use for test
test_query <- sprintf("SELECT * FROM %s", test_table)

# Connect to DSN string we were just given from environmental variable based on folder
oracle_odbc <- odbcConnect(odbc_dsn)

## Run test query
records_test <- sqlQuery(oracle_odbc, test_query)

## Display results of test query
records_test

```
